---
title: "Married and Unmarried Same and Different Sex Couples"
author: "James Chapman"
output: html_notebook
---

In this lab, you will be using the variables you made in Notebook 10. Run that file and make sure that the final `couples` data frame is in your working environment. All graphs must be publication quality with accurate titles. Don't forget to weight by `PERWT.x`.

```{r}
library(tidyverse)
library(viridis)
```


## Column graphs: Married and unmarried same- and different-sex couples
```{r fig.height = 10}
couples %>% group_by(YEAR, MARRIED, SSEX) %>% count(wt = PERWT.x) %>%
  ggplot(aes(YEAR, n, fill = MARRIED)) + 
  geom_col() + 
  facet_wrap(vars(SSEX), scales = "free_y") + 
  scale_x_continuous(breaks = c(1990, 2000)) + 
  scale_y_continuous(labels = scales::comma) + 
  theme_minimal(base_size = 20) + 
  labs(x = "Census Year", y = "Number of couples", fill = "Marital Status", title = "Number of  in 1990 and 2000")
```
## Line graphs: Allocation
```{r fig.height = 10}
couples %>% filter(MARRIED == "Unmarried") %>% 
  group_by(YEAR, SSEX) %>% count(ALLOCATED, wt = PERWT.x) %>% 
  mutate(TOTAL = sum(n), PCTALOC = n/TOTAL) %>% filter(ALLOCATED) %>%
  ggplot(aes(YEAR, PCTALOC, color = SSEX)) +
  geom_line() +
  theme_minimal() +
  scale_fill_brewer("Paired") +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = c(1990, 2000)) +
  labs(x = "Year", y = "Percent of Couples with Allocated Marital Status", title = "Percent of Couples with Allocated Marital Status per Year by Same Sex or Different Sex Couple", color = "Relationship Type")
```
## Population pyramids: Age and sex of person 1
```{r fig.height=10}
couples %>% mutate(SEX.x = factor(SEX.x, labels = c("Male", "Female"))) %>% filter(YEAR == 2000) %>%
  group_by(TYPE) %>% count(AGECAT, SEX.x, wt = PERWT.x) %>% 
  mutate(TOTAL = sum(n), n = ifelse(SEX.x == "Male", 0-n, n)) %>%
  ggplot(aes(AGECAT, n/TOTAL, fill = SEX.x)) + 
  geom_col() +
  coord_flip() + 
  #scale_fill_brewer(palette = "Set1") + 
  scale_y_continuous(breaks = seq(-.2, .2, .1), labels = c("20%", "10%","0%", "10%","20%")) +
  theme_minimal(base_size = 20) + 
  theme(legend.position = "bottom") +
  labs(x = "Percent of Category", y = "Age of Person 1", fill = "Sex of Person 1", title = "Age and Sex Distribution of Person 1 by Relationship Type") +
  facet_grid(cols = vars(TYPE))
```
## Box plots: Distribution of children among couples
```{r fig.height=10}
couples %>% filter(YEAR == 2000) %>%
  group_by(TYPE) %>% summarize(MIN = quantile(rep(NCHILD.x, PERWT.x), .1),
                           LOW = quantile(rep(NCHILD.x, PERWT.x), .25),
                           MED = quantile(rep(NCHILD.x, PERWT.x), .50),
                           HIGH = quantile(rep(NCHILD.x, PERWT.x), .75),
                           MAX = quantile(rep(NCHILD.x, PERWT.x), .90)) %>%
  ggplot(aes(x = factor(TYPE))) +
         geom_boxplot(aes(ymin = MIN, lower = LOW, middle = MED, upper = HIGH, ymax = MAX), stat = "identity") + 
         theme_minimal(base_size=20) +
         scale_fill_brewer(palette = "Paired") +
         labs(x = "Relationship Type", y = "Children in the Household", title = "Distribution of Children by Relationship Type (whiskers indicate 10th and 90th percentiles")
```
## Heat map: Primary language of Person 1
```{r fig.height = 10}
#First heat map:
couples %>% filter(YEAR == 2000) %>%
  group_by(TYPE) %>% count(LANG, wt = PERWT.x) %>% mutate(TOTAL = sum(n)) %>%
  ggplot(aes(x = TYPE, y = LANG, fill = n/TOTAL)) + geom_tile(color = "black") +
  scale_fill_gradient(low = "White", high = "Black", labels = scales::percent) +
  theme_minimal(base_size = 20) +
  labs(x = "Relationship Type", y = "Primary Language of Person 1", fill = "Percent of column", title = "Primary Language of Person 1 in a Relationship by Relationship Type")
```

